Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-04-15 main.Rmd
smart meter application update

### Levels

@@

We will discuss the effect of the mapping of cyclic granularities in this section. When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$-->. With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables. \autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarter-of-year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed. \autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.

@@

4. For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$. Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.

@@

R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here. While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively. Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year). Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year. Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed. If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped. Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month". Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Sections \ref{sec:cyclic-calendar}. <!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).  -->
Here, the index set $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ maps the half-hourly time instants to a set of positive integers. Then, multiple-order-up circular granularity $hour\_week$ and quasi-circular granularity $hour\_month$ could be computed from single-order-up granularities as follows:
<!-- $$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular}) -->
\begin{align}
  hour\_week(z) =  day\_week(z) + 24*hour\_day(z)
\end{align}

<!-- $$ day\_month(z) =   \lfloor z/2*24 \rfloor  - \sum_{w=0}^{k-1}\vert T_{w \mod 12*4} \vert$$ (\autoref{def:quasi-circular}) -->
\begin{align}
hour\_month(z) =  hour\_day(z) + 24*day\_month(z)
\end{align}

@@

From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior. Given the data structure <$C1$ $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet. Thus the problem is equivalent to taking 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualization. However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting any pairs together. \autoref{tab:harmony-tab} shows 13 harmonies pairs, each of which can be plotted together to look at the energy behavior from different perspectives.

>>>>

### Levels {#sec:levels}

@@

We will discuss the effect of the mapping of cyclic granularities in this section. When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets, then $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$--> With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables. \autoref{fig:allFig} (b) shows the letter value plot across weekday/weekend faceted by quarter-of-year and \autoref{fig:allFig} (c) shows the same two cyclic granularities with their mapping reversed. \autoref{fig:allFig} (b) helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig} (c) helps to compare quarters within weekend and weekday.

@@

4. For each $i \in \{1, 2, \dots, l\}$, the maximum pairwise distance could be chosen from  ${m \choose 2}$ distances using $D_{i} = max(d_{jk})/m(max(d_{jk}) - min(d_{jk}))$ $\forall j, k = {1, 2, \dots, m}$. Here, maximum is being adjusted against the range of the data to accommodate for comparison of maximum distances for different ranges.

@@

R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here. While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively. Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year). Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year. Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed relating two linear granularities at a time. If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units like hhour could be dropped. Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month". Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Section \ref{sec:cyclic-calendar}. <!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).  -->
with the index set mapping half-hourly time instants to a set of positive integers.

<!-- R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here. While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively. Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year). Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year. Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed. If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped. Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month". Now that we have a list of cyclic granularities to look at, we should be able to compute the multiple-order-up granularities using Sections \ref{sec:cyclic-calendar}. <!-- The lubridate package [@Grolemund2011-vm] provides tools to compute single-order-up granularities, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).  --> 
<!-- Here, the index set $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$ maps the half-hourly time instants to a set of positive integers. Then, multiple-order-up circular granularity $hour\_week$ and quasi-circular granularity $hour\_month$ could be computed from single-order-up granularities as follows: -->
<!-- <!-- $$ hour\_day(z) = \lfloor z/2 \rfloor \mod 7$$ (Definition \autoref{def:circular}) --> 
<!-- \begin{align} -->
<!--   hour\_week(z) =  day\_week(z) + 24*hour\_day(z) -->
<!-- \end{align} -->

<!-- <!-- $$ day\_month(z) =   \lfloor z/2*24 \rfloor  - \sum_{w=0}^{k-1}\vert T_{w \mod 12*4} \vert$$ (\autoref{def:quasi-circular}) --> 
<!-- \begin{align} -->
<!-- hour\_month(z) =  hour\_day(z) + 24*day\_month(z) -->
<!-- \end{align} -->

@@

From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior. Recalling the data structure <$C1$, $C2$, `general_supply_kwh`> that we would use for exploration, each of these six cyclic granularities can either be mapped to x-axis or to facet. Thus the problem is equivalent to choosing 2 granularities at a time from 6, which essentially is equivalent to having 30 data subsets for visualization. However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting pairs together. \autoref{tab:harmony-tab} shows 13 harmonies pairs with any cyclic granularities with levels more than 31 removed as effective exploration becomes difficult with many levels (Section \ref{sec:levels}). For each of \autoref{fig:bothcust} (b) and (c), $C1$ is the circular granularity day-type (weekday/ weekend) and $C2$ is hour of the day. The geometry used for displaying the distribution is chosen as area-quantiles \autoref{fig:bothcust} (b) and violins for \autoref{fig:bothcust} (b). \autoref{fig:bothcust} (a) displays reverse mapping of $C1$ and $C2$ with $C1$ now denoting hour of the day and $C2$ denoting day-type and geometry for displaying distribution is boxplots.

<!-- From the search list, we found six cyclic granularities for which we would like to derive insights of energy behavior. Given the data structure <$C1$, $C2$, `general_supply_kwh`>, each of those six cyclic granularities can either be mapped to x-axis or to facet. Thus the problem is equivalent to taking 2 granularities at a time from six, which essentially is equivalent to having 30 data subsets for visualization. However, harmony/clash pairs can be identified among those 30 possibilities to determine feasibility of plotting any pairs together. \autoref{tab:harmony-tab} shows 13 harmonies pairs, each of which can be plotted together to look at the energy behavior from different perspectives. -->
